Covariate analysis of HBCC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (DLPFC with CQN normlaisation)

### 2017-08-08 code/Rmd/HBCC_DLPFC_STAR_ENSEMBL_v75_CQN.Rmd
Minor sanity fix

### Differential expression analysis (ACC - DLPFC, SCZ - Control, Female - Male)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp2, fig.height=10, fig.width=6}
postAdjustCovars = c('Dx.Tissue', 'Gender', 'RIN', 'IntronicRate')

@@

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'CMC_ACC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('CMC_ACC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('CMC_ACC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_nCounts.tsv', sep = '\t', row.names=F, quote=F)
NCOUNT_OBJ = File('CMC_ACC_DLPFC_nCounts.tsv', name = 'Counts (estimated)', parentId = CODE$properties$id)
annotations(NCOUNT_OBJ) = annotations(CODE)
annotations(NCOUNT_OBJ)$dataSubType = 'estimatedCounts'
NCOUNT_OBJ = synStore(NCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('CMC_ACC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('CMC_ACC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DM1$design %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'CMC_ACC_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('CMC_ACC_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_ACC_DLPFC_netResidualExpression.tsv', 
                name = 'Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
annotations(nEXP_OBJ) = annotations(CODE)
annotations(nEXP_OBJ)$dataSubType = 'residualGeneExpForNetAnlz'
nEXP_OBJ = synStore(nEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store differential expression results
write.table(DE, file = 'CMC_ACC_DLPFC_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('CMC_ACC_DLPFC_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx.Tissue)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store enrichment results
write.table(enrichment.results, file = 'CMC_ACC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('CMC_ACC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r NCOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |

>>>>

### Differential expression analysis (DLPFC, SCZ - Control, Female - Male)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp2, fig.height=10, fig.width=6}
postAdjustCovars = c('Dx.Tissue', 'Gender', 'RIN', 'IntronicRate')

@@

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'HBCC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = File('HBCC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
annotations(COV_OBJ) = annotations(CODE)
annotations(COV_OBJ)$dataSubType = 'covariates'
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'HBCC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('HBCC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'HBCC_DLPFC_nCounts.tsv', sep = '\t', row.names=F, quote=F)
NCOUNT_OBJ = File('HBCC_DLPFC_nCounts.tsv', name = 'Counts (estimated)', parentId = CODE$properties$id)
annotations(NCOUNT_OBJ) = annotations(CODE)
annotations(NCOUNT_OBJ)$dataSubType = 'estimatedCounts'
NCOUNT_OBJ = synStore(NCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'HBCC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('HBCC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLCPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'HBCC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('HBCC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
annotations(OFFSET_OBJ) = annotations(CODE)
annotations(OFFSET_OBJ)$dataSubType = 'offset'
OFFSET_OBJ = synStore(OFFSET_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DM1$design %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'HBCC_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('HBCC_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'HBCC_DLPFC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('HBCC_DLPFC_netResidualExpression.tsv', 
                name = 'Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
annotations(nEXP_OBJ) = annotations(CODE)
annotations(nEXP_OBJ)$dataSubType = 'residualGeneExpForNetAnlz'
nEXP_OBJ = synStore(nEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store differential expression results
write.table(DE, file = 'HBCC_DLPFC_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('HBCC_DLPFC_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx.Tissue)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store enrichment results
write.table(enrichment.results, file = 'HBCC_DLPFC_EnrichResults.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('HBCC_DLPFC_EnrichResults.tsv', 
                  name = 'Enrichment Analysis Results (Dx.Tissue)', 
                  parentId = CODE$properties$id)
annotations(ENRICH_OBJ) = annotations(CODE)
annotations(ENRICH_OBJ)$dataSubType = 'enrichResults'
ENRICH_OBJ = synStore(ENRICH_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r NCOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Enrichment Analysis                        |  `r ENRICH_OBJ$properties$id`    |
